---
title: "upper beaver"
author: "Claire Tortorelli"
date: "January 2, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vegan)
library(expss)
library(tidyverse)
library(here)
library(lme4)
#install.packages(c("wordcloud","tm"),repos="http://cran.r-project.org")
library(wordcloud)
library(tm)
library(emmeans)
```

```{r read in data}

ub_veg <- read.csv(here("raw-data/ubdata_16-20_species_average.csv"))

sp <- read.csv(here("raw-data/speciesList_USDAplants.csv"))
```

## Prep data for analysis

Calculate functional group cover for overlay

```{r}
#convert to long format
df <- ub_veg %>% gather(key = "CT_CODE", value = "cover", -(1:5))

#merge with functional groups

func <- merge(df, sp)

#check to make sure all names match
# spname <- colnames(ub_veg)
# spname %in% sp$CT_CODE #looks good!


#add cover by plot for each functional group

func_sum <- func %>% group_by(PLOT_YR, functionalGroup) %>%
              summarise(funcSum = sum(cover)) %>%
              spread(., functionalGroup, funcSum)

funcdf <- merge(func_sum, ub_veg[,1:5])

```

Compare diversity!
```{r}
div <- ub_veg

div$richness <- rowSums(div[,6:114] != 0)
div$shan <- diversity(div[,6:114], index = "shannon")


div$YEAR <- factor(div$YEAR)
#merge unburned and control unburned plots
div$BURN2 <- div$BURN17
div$BURN2[which(div$BURN2 == "c_unburned")] <- "unburned"

div$BURN17 <- factor(div$BURN17, levels = c("unburned", "<50", ">50"))
```


model change since fire with pre-fire ventenata and brome as baseline
```{R}
#separate pre-fire data from post-fire data

div_postfire <- subset(div, YEAR != '16')
div_prefire <- subset(div, YEAR == '16')


div_prefire$richness_prefire <- div_prefire$richness
div_prefire$shan_prefire <- div_prefire$shan


div_prefiresub <- div_prefire[c("PLOT", "richness_prefire", "shan_prefire")]

#merge prefire and post fire dataframes
div_merge <- merge(div, div_prefiresub, by = "PLOT")

```

Model diversity
```{r model diversity}

mshan <- lmer(shan ~ YEAR*BURN17 + shan_prefire + (1|PLOT), data = div_merge)
plot(mshan)

summary(mshan)
confint(mshan)

mrich <- lmer(richness ~ YEAR*BURN17 + richness_prefire + (1|PLOT), data = div_merge)
plot(mrich)

summary(mrich)
confint(mrich)


(emmeans(mrich, specs = revpairwise ~ BURN17 | YEAR, adjust = NULL))
(emmeans(mshan, specs = revpairwise ~ BURN17 | YEAR, adjust = NULL))


library(car)
Anova(mrich, type = 2)
Anova(mshan, type = 2)
```


change in functional group - shrubs?
```{r}
funcdf20 <- funcdf %>% subset(YEAR == 20) 
funcdfburn <- funcdf %>% subset(BURN17 == ">50")

plot(funcdfburn$Shrub ~ factor(funcdfburn$YEAR))

``` 

--------------------------------------------------

# Community NMS

remove overstory species
```{r}
# tree <- c("PIPO", "PSME", "JUOC")
# 
# ub_vegund <- subset(ub_veg, select = -c(PIPO, PSME, JUOC))

```



remove rare species (occur in fewer than 5% of plots)
```{r rare species}
# 0.03*148 = 4.44
# 0.05*148 = 7.4

#subset df so it just consists of species data
sp <- ub_veg[,6:114] #without canopy

#create a vector that counts number of rows > 0
count <- apply(sp, 2, function(x){sum(x > 0)})
# remove species that occured in fewer than 8 plots
sp_sub <- sp[,-which(count < 8)]

ub_vegnorare <- cbind(ub_veg[,1:5], sp_sub)
```

removed rare species that occurred in fewer than 5% of plots leaving 49 species from the original 115

## NMS with all years

```{r ordintions}
sp4nms <- ub_vegnorare[,c(1,6:54)]


#set plot to row names 
rownames(sp4nms) <- sp4nms$PLOT_YR 
sp4nms$PLOT_YR  <- NULL 


nms <- metaMDS(sqrt(sp4nms), distance='bray', k=2, trymax=50, autotransform=FALSE, wasscores=TRUE, noshare=FALSE) 
```


```{r}
#convert year to factor
ub_vegnorare$YEAR <- as.factor(ub_vegnorare$YEAR)
ub_vegnorare$PLOT <- as.factor(ub_vegnorare$PLOT)
```

```{r}
envdf1 <- funcdf[,2:10]

#fit to ordination
env.fit <- envfit(nms, envdf1, perm=500)
```


plot
```{r}
plot(nms, type = "n")

#ordiarrows (nms, groups = ub_vegnorare$PLOT, order.by = ub_vegnorare$YEAR, label = FALSE, length = .1, display = "sites")

# ordiellipse(nms, ub_vegnorare$BURN17, col = 1:4, kind="sd", label = TRUE)
ordiellipse(nms, ub_vegnorare$YEAR, col = 1:4, kind="sd", label = TRUE, show.groups = c(16, 18, 19, 20))


# with(ub_vegnorare, points(nms, display = "sites", col = c("#606c38", "#dda15e", "#bc6c25")[YEAR], cex=1.25))
with(ub_vegnorare, points(nms, display = "species"))

plot(env.fit)
```

### NMS with only YR 2016 (pre-fire) & 2020 (3 yrs post fire)
```{R with only 16 and 20 data}

sp4nms2 <- ub_vegnorare[which(ub_vegnorare$YEAR == "16" | ub_vegnorare$YEAR == "20"),c(1,6:54)]
#set plot to row names 
rownames(sp4nms2) <- sp4nms2$PLOT_YR 
sp4nms2$PLOT_YR  <- NULL 

nms2 <- metaMDS(sqrt(sp4nms2), distance='bray', k=2, trymax=50, autotransform=FALSE, wasscores=TRUE, noshare=FALSE) 
```
```{r}
#rotate to weigh invasive annual grasses on x axis
nms2_rot <- with(funcdf_sub, MDSrotate(nms2, Invasive.Annual.Grass))
```



```{r}
ub_vegnorare2 <- ub_vegnorare[which(ub_vegnorare$YEAR == 16 | ub_vegnorare$YEAR == 20),]

ub_vegnorare2$BURN2 <- ub_vegnorare2$BURN17
ub_vegnorare2$BURN2[which(ub_vegnorare2$BURN2 == "c_unburned")] <- "unburned"

ub_vegnorare2$BURN2_YR <- paste(ub_vegnorare2$BURN2, ub_vegnorare2$YEAR, sep = "_")
ub_vegnorare2$BURN17_YR <- paste(ub_vegnorare2$BURN17, ub_vegnorare2$YEAR, sep = "_")

ub_vegnorare2$BURN2_YR <- factor(ub_vegnorare2$BURN2_YR, levels = c("unburned_16", "<50_16", ">50_16", "unburned_20", "<50_20", ">50_20"))

ub_vegnorare2$BURN2 <- factor(ub_vegnorare2$BURN2, levels = c("unburned", "<50", ">50"))
ub_vegnorare2$YEAR <- factor(ub_vegnorare2$YEAR)
```

```{r}
#extract scores to plot with vectors
speciesScores <- nms2_rot$species
siteScores <- nms2_rot$points

siteScores_burn <- cbind(data.frame(siteScores), ub_vegnorare2[,"BURN17"])

species <- rownames(speciesScores)


```





```{r fig.height=6, fig.width=11}
svg("speciesNMS.svg", width = 11, height = 6)

par(mfrow=c(1,2))

plot(nms2_rot, type = "n")

with(ub_vegnorare2, points(nms2_rot, display = "sites", col = c("#52796f", "#dda15e", "#bc6c25")[BURN2], cex=1.25, pch = c(16, 17)[YEAR]))

#add species text
# text(nms2_rot, display = "spec", cex=0.7, col="grey")

ordiellipse(nms2_rot, ub_vegnorare2$BURN2_YR,
            scaling = 3, label = FALSE, draw = "polygon",
            col = c("#52796f", "#dda15e", "#bc6c25"),
            kind = "sd",
            alpha = 40,
            show.groups = c("unburned_16", "<50_16", ">50_16"),
            border = "white")



#add the linear vectors for functional groups
plot(env.fit, col = "black", font = 2, cex=0.8)

#--------------------------------------------------------------------

# plot post-fire (same NMS with 2020 ellipses)

plot(nms2_rot, type = "n")


with(ub_vegnorare2, points(nms2_rot, display = "sites", col = c("#52796f", "#dda15e", "#bc6c25")[BURN2], cex=1.25, pch = c(16, 17)[YEAR]))

#with(ub_vegnorare2, points(nms2_rot, display = "species", col = "grey", cex=1.25, pch = 4), label = TRUE)
# text(nms2_rot, display = "spec", cex=0.7, col="grey")

ordiellipse(nms2_rot, ub_vegnorare2$BURN2_YR,
            scaling = 3, label = FALSE, draw = "polygon",
            col = c("#52796f", "#dda15e", "#bc6c25"),
            kind = "sd",
            alpha = 40,
            show.groups = c("unburned_20", "<50_20", ">50_20"),
            border = "white")

plot(env.fit, col = "black", font = 2, cex=0.8)



dev.off()
```


## NMS with only YR 2016 (pre-fire) & 2020 (3 yrs post fire) by functional group
```{R with only 16 and 20 data}
funcdf$BURN2 <- funcdf$BURN17
funcdf$BURN2[which(funcdf$BURN2 == "c_unburned")] <- "unburned"

funcdf$BURN2_YR <- paste(funcdf$BURN2, funcdf$YEAR, sep = "_")
funcdf$BURN17_YR <- paste(funcdf$BURN17, funcdf$YEAR, sep = "_")

#remove canopy "tree"
funcdf2 <- funcdf[,-11]

#subset dataframe for pre and post fire data
sp4nmsfunc <- funcdf2[which(funcdf2$YEAR == 16 | funcdf2$YEAR == 20), 1:10]



#set plot to row names 
rownames(sp4nmsfunc) <- sp4nmsfunc$PLOT_YR 
sp4nmsfunc$PLOT_YR  <- NULL 
```

```{r}
nmsfunc <- metaMDS(sqrt(sp4nmsfunc), distance='bray', k=2, trymax=500, autotransform=FALSE, wasscores=TRUE, noshare=FALSE) 

```

```{r}
funcdf_sub <- funcdf2[which(funcdf$YEAR == 16 | funcdf2$YEAR == 20),]

funcdf_sub$YEAR <- factor(funcdf_sub$YEAR)
funcdf_sub$BURN17 <- factor(funcdf_sub$BURN17, levels = c("c_unburned", "unburned", "<50", ">50"))
funcdf_sub$BURN2 <- factor(funcdf_sub$BURN2, levels = c("unburned", "<50", ">50"))

#add factor levles
funcdf_sub$BURN17_YR <- factor(funcdf_sub$BURN17_YR, levels = c("c_unburned_16","unburned_16", "<50_16", ">50_16", "c_unburned_20","unburned_20", "<50_20", ">50_20"))

funcdf_sub$BURN2_YR <- factor(funcdf_sub$BURN2_YR, levels = c("unburned_16", "<50_16", ">50_16", "unburned_20", "<50_20", ">50_20"))
```


```{r}
#rotate to weigh invasive annual grasses on x axis
nmsfunc_rot <- with(funcdf_sub, MDSrotate(nmsfunc, Invasive.Annual.Grass))
```


```{r}
#extract scores to plot with vectors

speciesScoresFunc <- nmsfunc_rot$species
siteScoresFunc <- nmsfunc_rot$points

siteScores_burnFunc <- cbind(data.frame(siteScoresFunc), funcdf_sub[, c("BURN2", "YEAR")])

speciesFunc <- rownames(speciesScoresFunc)
```

```{r}
envdf <- funcdf_sub[,2:10]

#fit to ordination
env.fit <- envfit(nmsfunc_rot, envdf, perm=500)
```



plot nms functional groups
```{r fig.height=6, fig.width=11}
#svg("functionalGroupNMS.svg", width = 11, height = 6)

par(mfrow=c(1,2))

plot(nmsfunc_rot, type = "n")

with(funcdf_sub, points(nmsfunc_rot, display = "sites", col = c("#52796f", "#dda15e", "#bc6c25")[BURN2], cex=1.25, pch = c(16, 17)[YEAR]))



ordiellipse(nmsfunc_rot, funcdf_sub$BURN2_YR,
            scaling = 3, label = FALSE, draw = "polygon",
            col = c("#52796f", "#dda15e", "#bc6c25"),
            kind = "sd",
            alpha = 40,
            show.groups = c("unburned_16", "<50_16", ">50_16"),
            border = "white")

#add the linear vectors for functional groups
plot(env.fit, col = "black", font = 2, cex=0.8)

#--------------------------------------------------------------------

# plot post-fire (same NMS with 2020 ellipses)

plot(nmsfunc_rot, type = "n")


with(funcdf_sub, points(nmsfunc_rot, display = "sites", col = c("#52796f", "#dda15e", "#bc6c25")[BURN2], cex=1.25, pch = c(16, 17)[YEAR]))



ordiellipse(nmsfunc_rot, funcdf_sub$BURN2_YR,
            scaling = 3, label = FALSE, draw = "polygon",
            col = c("#52796f", "#dda15e", "#bc6c25"),
            kind = "sd",
            alpha = 40,
            show.groups = c("unburned_20", "<50_20", ">50_20"),
            border = "white")

#add the linear vectors for functional groups
plot(env.fit, col = "black", font = 2, cex=0.8)

# arrows(0, 0, speciesScoresFunc[, 1], speciesScoresFunc[, 2], length=0.1, angle=20, col ="black")

#set text
# text(speciesScoresFunc[, 1], speciesScoresFunc[, 2], pos = 3, offset = 0.5, col = "black", font = 2, rownames(speciesScoresFunc), cex=0.8)

#arrows(0, 0, variableScores[, 1], variableScores[, 2], length=0.1, angle=20, col ="grey")

#add species points
#points(speciesScoresFunc[, 1], speciesScoresFunc[, 2], pch=15, cex=0.75, col="grey")

# #add ordi-arrows from pre to post fire
# ordiarrows (nmsfunc_rot, groups = funcdf_sub$PLOT, order.by = funcdf_sub$YEAR,
#             label = FALSE,
#             length = .1,
#             display = "sites", col = "grey")

#dev.off()
```

```{r fig.height=6, fig.width=11}
#svg("functionalGroupNMS.svg", width = 11, height = 6)

par(mfrow=c(1,2))

plot(nmsfunc_rot, type = "n")

with(funcdf_sub, points(nmsfunc_rot, display = "sites", col = c("#1C1C5D", "#52796f", "#dda15e", "#bc6c25")[BURN17], cex=1.25, pch = c(16, 17)[YEAR]))



ordiellipse(nmsfunc_rot, funcdf_sub$BURN17_YR,
            scaling = 3, label = FALSE, draw = "polygon",
            col = c("#1C1C5D", "#52796f", "#dda15e", "#bc6c25"),
            kind = "sd",
            alpha = 40,
            show.groups = c("c_unburned_16","unburned_16", "<50_16", ">50_16"),
            border = "white")

#add the linear vectors for functional groups
plot(env.fit, col = "black", font = 2, cex=0.8)

#--------------------------------------------------------------------

# plot post-fire (same NMS with 2020 ellipses)

plot(nmsfunc_rot, type = "n")


with(funcdf_sub, points(nmsfunc_rot, display = "sites", col = c("#1C1C5D", "#52796f", "#dda15e", "#bc6c25")[BURN17], cex=1.25, pch = c(16, 17)[YEAR]))



ordiellipse(nmsfunc_rot, funcdf_sub$BURN17_YR,
            scaling = 3, label = FALSE, draw = "polygon",
            col = c("#1C1C5D", "#52796f", "#dda15e", "#bc6c25"),
            kind = "sd",
            alpha = 40,
            show.groups = c("c_unburned_20","unburned_20", "<50_20", ">50_20"),
            border = c("#1C1C5D", "#52796f", "#dda15e", "#bc6c25"))

#add the linear vectors for functional groups
plot(env.fit, col = "black", font = 2, cex=0.8)
```



Compare groups with MRPP
```{r}
sp4nmsfunc16 <- sp4nmsfunc[1:37,]
sp4nmsfunc20 <- sp4nmsfunc[38:74,]

func.mrpp16 <- with(funcdf_sub[which(funcdf_sub$YEAR == 16),], mrpp(sp4nmsfunc16, BURN2, distance = "bray"))
func.mrpp16

func.mrpp20 <- with(funcdf_sub[which(funcdf_sub$YEAR == 20),], mrpp(sp4nmsfunc20, BURN2, distance = "bray"))
func.mrpp20


#compare pre and post treatment functional composition
funcdf_unburned <- funcdf_sub %>% filter(BURN2 == "unburned") %>% column_to_rownames(var = "PLOT_YR")

funcdf_lowsev <- funcdf_sub %>% filter(BURN2 == "<50") %>% column_to_rownames(var = "PLOT_YR")

funcdf_highsev <- funcdf_sub %>% filter(BURN2 == ">50") %>% column_to_rownames(var = "PLOT_YR")


(funcmp.unburn <- with(funcdf_unburned, mrpp(funcdf_unburned[,2:9], YEAR, distance = "bray")))
(funcmp.lowsev <- with(funcdf_lowsev, mrpp(funcdf_lowsev[,2:9], YEAR, distance = "bray")))
(funcmp.highsev <- with(funcdf_highsev, mrpp(funcdf_highsev[,2:9], YEAR, distance = "bray")))



#explore means
func.md16 <- with(funcdf_sub[which(funcdf_sub$YEAR == 16),], meandist(vegdist(sp4nmsfunc16), BURN2, distance = "bray"))
summary(func.md16)

plot(func.md16, kind = "histogram")

func.md20 <- with(funcdf_sub[which(funcdf_sub$YEAR == 20),], meandist(vegdist(sp4nmsfunc20), BURN2, distance = "bray"))
summary(func.md20)

plot(func.md20, kind = "histogram")
```

Compare unburned_control and unburned plots pre-fire
```{r}
(funcmp.unburn <- with(funcdf_unburned[1:21,], mrpp(funcdf_unburned[1:21,2:9], BURN17, distance = "bray")))

```


plot with ggplot
```{r}
# siteScores.df <- siteScores_burnFunc
# siteScores.df$plot_yr <- rownames(siteScores.df)
# siteScores.df$YEAR <- factor(siteScores.df$YEAR)
# 
# 
# speciesScores.df <- data.frame(speciesScoresFunc)
# speciesScores.df$species <- rownames(speciesScores.df)
# head(speciesScores.df)
# ```
# 
# ```{r fig.height=6, fig.width=6}
# library(ggrepel)
# library(ggordiplots)
# ggplot() +
#   # gg_ordiplot(nmsfunc,
#   #             groups = c(funcdf_sub$BURN2, funcdf_sub$YEAR),
#   #             kind = "sd")
#   geom_point(data=speciesScores.df,aes(x=MDS1,y=MDS2), col = "grey", size=2, pch = 17) + # add the point markers
#   geom_point(data=siteScores.df,aes(x=MDS1,y=MDS2, colour=BURN2, shape = YEAR),size=3) + # add the point markers
#   geom_text_repel(data=speciesScores.df, aes(x=MDS1,y=MDS2,label=species), alpha=0.5, size = 3, force = 1) + # add the species labels
#   #geom_text(data=data.scores,aes(x=NMDS1,y=NMDS2,label=site),size=6,vjust=0) +  # add the site labels
#   scale_colour_manual(values=c("unburned" = "#606c38", "<50" = "#dda15e", ">50" = "#bc6c25")) +
#   coord_equal() +
#   theme_bw(base_size = 13) +
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
# 
# 
# 

```



Calculate trajectories
```{r}
library(vegclust)
library(RColorBrewer)
library(smacof)
#calculate distnace matrix for nms
nmsdist <- as.matrix(metaMDSdist(sp4nms, distance='bray', k=2, trymax=50, autotransform=FALSE, wasscores=TRUE, noshare=FALSE))
nmsdist2 <- as.matrix(metaMDSdist(sp4nms2, distance='bray', k=2, trymax=50, autotransform=FALSE, wasscores=TRUE, noshare=FALSE))

```

```{r}
tlen <- trajectoryLengths(nmsdist, ub_vegnorare$PLOT, ub_vegnorare$YEAR)
tlen2 <- trajectoryLengths(nmsdist2, ub_vegnorare2$PLOT, ub_vegnorare2$YEAR)
# tdist <- trajectoryDistances(nmsdist, ub_vegnorare$PLOT, ub_vegnorare$YEAR, distance.type = "Hausdorff")

trajectoryPlot(nmsdist, ub_vegnorare$PLOT, ub_vegnorare$YEAR, traj.colors = ub_vegnorare$IAG.preburn_BURN)

trajectoryPlot(nmsdist2, ub_vegnorare2$PLOT, ub_vegnorare2$YEAR, traj.colors = ub_vegnorare2$IAG.preburn_BURN, distance.type = "Hausdorff")
```