---
title: "upper beaver"
author: "Claire Tortorelli"
date: "January 2, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(MASS)
library(nlme)
library(emmeans)
library(tidyverse)
library(sjPlot)
library(car)
```

```{r read in data}

ubveg_iag <- read.csv(here("raw-data", "ubdata_16-20_species_tally.csv"))
ubveg <- read.csv(here("raw-data/ubdata_16-20_species_average.csv"))
```

```{r}

#change year to factor
ubveg_iag$YEAR <- as.factor(ubveg_iag$YEAR)
```



```{r}
#combine control unburned and uburned plots for one unburned category
ubveg_iag[ubveg_iag=="c_unburned"] <- "unburned"

#redefine levels for burn17
ubveg_iag$BURN17 <- factor(ubveg_iag$BURN17,
levels = c("unburned", "<50", ">50") )

```

model change since fire with pre-fire ventenata and brome as baseline
```{R}
#separate pre-fire data from post-fire data

ubiag_postfire <- subset(ubveg_iag, YEAR != '16')
ubiag_prefire <- subset(ubveg_iag, YEAR == '16')


ubiag_prefire$VEDU80_prefire <- ubiag_prefire$VEDU80
ubiag_prefire$BRTE80_prefire <- ubiag_prefire$BRTE80
ubiag_prefire$BRJA80_prefire <- ubiag_prefire$BRJA80

ubiag_prefiresub <- ubiag_prefire[c("PLOT", "VEDU80_prefire", "BRTE80_prefire", "BRJA80_prefire")]

#merge prefire and post fire dataframes
ubiag_merge <- merge(ubiag_postfire, ubiag_prefiresub, by = "PLOT")

```

Explore data - how does pre-fire cover influence post-fire invasion?
```{r}
df20 <- filter(ubiag_merge, YEAR == '20')
df19 <- filter(ubiag_merge, YEAR == '19')
plot(df20$VEDU80 ~ df20$VEDU80_prefire, col = df20$BURN17)

lmvd <- lm(VEDU80 ~ VEDU80_prefire, data = df20)
lmbt <- lm(BRTE80 ~ BRTE80_prefire, data = df20)
lmbj <- lm(BRJA80 ~ BRJA80_prefire, data = df20)
# 
# summary(lmvd)
# summary(lmbt)
# summary(lmbj)
# 
# plot(lmvd)
# plot(lmbt)
# plot(lmbj)
```


## Model IAG response to burning*year

model IAGs separately using binomial generalized linear mixed models (logistic regression)
```{r}
library(lme4)
#https://aosmith.rbind.io/2020/08/20/simulate-binomial-glmm/

m.vd <- glmer(cbind(VEDU80, 80 - VEDU80) ~ BURN17*YEAR + VEDU80_prefire + (1 | PLOT), data = ubiag_merge, family = binomial)

m.bt <- glmer(cbind(BRTE80, 80 - BRTE80) ~ BURN17*YEAR + BRTE80_prefire +(1 | PLOT), data = ubiag_merge, family = binomial)
m.bt2 <- update(m.bt, control = glmerControl(optimizer="bobyqa")) ##trouble converging..fit with a different optimizer 

m.bj <- glmer(cbind(BRJA80, 80 - BRJA80) ~ BURN17*YEAR + BRJA80_prefire +(1 | PLOT), data = ubiag_merge, family = binomial)
m.bj2 <- update(m.bj, control = glmerControl(optimizer="bobyqa"))

# summary(m.bj)

#check for overdispersion - f the ratio considerably larger than 1, then it indicates that we have an overdispersion issue.
deviance(m.vd)/df.residual(m.vd) 
deviance(m.bt2)/df.residual(m.bt2) 
deviance(m.bj2)/df.residual(m.bj2) 

#looks good!
```

explore model results
```{r}
#examine results
# summary(m.vd)
emmeans(m.vd, revpairwise ~ BURN17 + VEDU80_prefire, adjust = NULL)
emvd <- emmeans(m.vd, trt.vs.ctrl ~ BURN17 | YEAR + VEDU80_prefire, adjust = NULL)
Anova(m.vd)

# summary(m.bt2)
embt <- emmeans(m.bt2, trt.vs.ctrl ~ BURN17 | YEAR + BRTE80_prefire, adjust = NULL)
Anova(m.bt2)

# summary(m.bj2)
embj <- emmeans(m.bj2, trt.vs.ctrl ~ BURN17 | YEAR + BRJA80_prefire, adjust = NULL)
Anova(m.bj2)

```

### Plot contrasts
```{r}
#extract contrasts to table for each species
emvd2 <- emvd$contrasts %>%
     confint() %>% rbind()
embt2 <- embt$contrasts %>%
     confint() %>% rbind()
embj2 <- embj$contrasts %>%
     confint() %>% rbind()

levels(emvd2$contrast) <- list("low severity - unburned" = "<50 - unburned", "high severity - unburned" = ">50 - unburned")
levels(emvd2$YEAR) <- list("2018" = "18", "2019" = "19", "2020" = "20")

levels(embt2$contrast) <- list("low severity - unburned" = "<50 - unburned", "high severity - unburned" = ">50 - unburned")
levels(embt2$YEAR) <- list("2018" = "18", "2019" = "19", "2020" = "20")

levels(embj2$contrast) <- list("low severity - unburned" = "<50 - unburned", "high severity - unburned" = ">50 - unburned")
levels(embj2$YEAR) <- list("2018" = "18", "2019" = "19", "2020" = "20")

```

contrasts plotting function
```{r}
plotcontrast <- function(df){
 p <-  ggplot(df, aes(x = YEAR, y = estimate, group = contrast, color = contrast) ) + 
		geom_point(size = 3.2, position = position_dodge(0.2)) + # Add points
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = .5, alpha = 0.5, size = 1, position = position_dodge(0.2)) +
		labs(x = NULL, y = "difference in log-odds") + # Label axes
		theme_bw(base_size = 15) + # Make black and white,
								# increase base text size
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
	geom_hline(yintercept = 0, colour = "dark grey", lty = 2) +
	scale_color_manual(values = c("grey", "black"))+
  theme(legend.title = element_blank()) +
  ylim(min = -5, max = 3)

 return(p)
}


```

plot contrasts
```{r}
vcontrast <- plotcontrast(emvd2) +  theme(legend.position = "bottom") + ggtitle("")
#pre-fire VD hits 5.76 = 7.2% cover

btcontrast <- plotcontrast(embt2) +  theme(legend.position = "NULL") + ylab("") + ggtitle("")
#pre-fire bt hits - 4.05 = 5.06 % cover

bjcontrast <- plotcontrast(embj2) +  theme(legend.position = "NULL") + ylab("") + ggtitle("")
#prefire bj hits - 3.11 = 3.89 % cover

```

### Plot raw data
```{r}
#plot raw data with standard errors
#summarize data
library(Rmisc)
library(cowplot)
library(ggpubr)

ubveg_iag$VEDU_perc <- ubveg_iag$VEDU80/80*100
ubveg_iag$BRTE_perc <- ubveg_iag$BRTE80/80*100
ubveg_iag$BRJA_perc <- ubveg_iag$BRJA80/80*100

ubveg_iag$YEAR2 <- ubveg_iag$YEAR
ubveg_iag$BURN2 <- ubveg_iag$BURN17
levels(ubveg_iag$BURN2) <- list("unburned" = "unburned", "low severity" = "<50", "high severity" = ">50")
levels(ubveg_iag$YEAR2) <- list("2016" = "16", "2018" = "18", "2019" = "19", "2020" = "20")
#ventenata
vdsum <- summarySE(ubveg_iag, measurevar="VEDU_perc", groupvars=c("BURN2","YEAR2"))
#brte
btsum <- summarySE(ubveg_iag, measurevar="BRTE_perc", groupvars=c("BURN2","YEAR2"))
#brja
bjsum <- summarySE(ubveg_iag, measurevar="BRJA_perc", groupvars=c("BURN2","YEAR2"))


```

raw data plotting function
```{r fig.height=4.25, fig.width=8.5}
plotIAG <- function(df,y.var){
 p <-  ggplot(df, aes(x = as.integer(as.character(YEAR2)), y = y.var, group = BURN2, color = BURN2) ) + 
		geom_line(position = position_dodge(0.2))+
		geom_point(size = 3.2, position = position_dodge(0.2)) + # Add points
  geom_errorbar(aes(ymin = y.var - se, ymax = y.var + se), width = .5, alpha = 0.5, size = 1, position = position_dodge(0.2))+
		labs(x = NULL, y = "Mean cover(%)") + # Label axes
		theme_bw(base_size = 15) + # Make black and white,
								# increase base text size
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
	
	scale_color_manual(values = c("#52796f", "#dda15e", "#bc6c25"))+
  theme(legend.title = element_blank()) +
  geom_vline(xintercept = 2017, colour = "dark grey", lty = 2)+
  ylim(min = 0, max = 35)

 return(p)
}
```

plot raw data
```{r fig.height=4, fig.width=9}

rawvd <- plotIAG(vdsum, vdsum$VEDU_perc) + ggtitle(expression(italic("V. dubia"))) +  theme(legend.position = "bottom")

rawbt <- plotIAG(btsum, btsum$BRTE_perc) + ggtitle(expression(italic("B. tectorum"))) + theme(legend.position = "NULL") + ylab("")

rawbj <- plotIAG(bjsum, bjsum$BRJA_perc) + ggtitle(expression(italic("B. arvensis"))) + theme(legend.position = "NULL") + ylab("")
```

-------------------------------------------------------------------------

## Part 2: Model proportion IAG compared to total understory veg (After fire)

Organize species data
```{r}
#convert abudnance to number of hits for modeling so that all species are on the same scale (hits/40)
df <- ubveg

ub_veg <- df[6:114] * 40/100
ub_veg <- cbind(df[1:5], ub_veg)

#add  IAG col
ub_veg$IAG <- (ub_veg$VEDU + ub_veg$BRJA + ub_veg$BRTE)

#add CANOPY col
ub_veg$CANOPY <- ub_veg$PIPO + ub_veg$PSME + ub_veg$JUOC

#add all veg col
ub_veg$all_veg <- rowSums(ub_veg[,6:114])

#non IAG col
ub_veg$all_veg_woIAG <- ub_veg$all_veg - ub_veg$IAG

#all veg without the canopy and without IAG
ub_veg$veg.noiag.nocanopy <- ub_veg$all_veg_woIAG - ub_veg$CANOPY

#all veg without canopy cover
ub_veg$all_veg.nocanopy <- ub_veg$all_veg - ub_veg$CANOPY

#add proportion IAG col (out of understory veg)
ub_veg$prop.IAG <- ub_veg$IAG/ub_veg$all_veg.nocanopy
```

Organize non-veg factors
```{r}
#change year to factor
ub_veg$YEAR <- as.factor(ub_veg$YEAR)

# combine unburned and c_unburned 
ub_veg$burn2 <- ub_veg$BURN17
ub_veg$burn2[ub_veg$burn2=="c_unburned"] <- "unburned"

#redefine levels for burn17
ub_veg$burn2 <- factor(ub_veg$burn2,
levels = c("unburned", "<50", ">50") )
```


```{R}
#organize data so that prefire IG cover has its own column

ubiag_postfire <- subset(ub_veg, YEAR != '16')
ubiag_prefire <- subset(ub_veg, YEAR == '16')


ubiag_prefire$IAG_prefire <- ubiag_prefire$IAG
ubiag_prefire$all_veg.nocanopy_prefire <- ubiag_prefire$all_veg.nocanopy

ubiag_prefiresub <- ubiag_prefire[c("PLOT", "IAG_prefire", "all_veg.nocanopy_prefire")]

#merge prefire and post fire dataframes
ub_merge <- merge(ubiag_postfire, ubiag_prefiresub, by = "PLOT")
```


model proportion of IAG to total non canopy vegetation
```{r}
hist(ub_merge$prop.IAG) 

qplot(burn2, prop.IAG, data = ub_veg)

# give our response as both the number of successes and failures (IAG and all_veg-IAG)

gm1 <- glmer(cbind(IAG, all_veg.nocanopy - IAG) ~ burn2*YEAR + IAG_prefire + (1 | PLOT),
              data = ub_merge, family = binomial)

#fit with a different optimizer to allow convergence
ss <- getME(gm1,c("theta","fixef"))
gm2 <- update(gm1,start=ss,control=glmerControl(optimizer="bobyqa",
                            optCtrl=list(maxfun=2e5)))

# check for overdispersion

# qplot(fitted(gm2), resid(gm2) ) +
# theme_bw() +
# labs(x = "Fitted",
# y = "Deviance Residuals") 

deviance(gm2)/df.residual(gm2) 
#low overdispersion

# summary(gm2)

#Wald Chi-square for interaction effect
Anova(gm1)
```
strong effect of pre-fire IAG, positive effect of year 20 and some negative effect of high severity. No evidence of interaction, although negative effect of burning seems to decrease with time after fire

get contrasts
```{r contrasts IAG prop}
#use trt.vs.ctrl to compare all groups to reference level (unburned)
emiag <- emmeans(gm1, specs = trt.vs.ctrl ~ burn2|YEAR  + IAG_prefire, adjust = NULL)

emiag2 <- emiag$contrasts %>%
     confint() %>% rbind()

levels(emiag2$contrast) <- list("low severity - unburned" = "<50 - unburned", "high severity - unburned" = ">50 - unburned")
levels(emiag2$YEAR) <- list("2018" = "18", "2019" = "19", "2020" = "20")
```

plot contrasts
```{r plot contrasts IAG prop}
iagcontrast <- plotcontrast(emiag2) +  theme(legend.position = "NULL") + ylab("") + ggtitle("")
#prefire iag hits (total) - 6.46 = 8.11 % cover

```

plot raw data
```{r fig.height=4, fig.width=3.5}
#plot raw data with standard errors
#summarize data
library(Rmisc)
#ventenata
iagsum <- summarySE(ub_veg, measurevar="prop.IAG", groupvars=c("burn2","YEAR"))
levels(iagsum$burn2) <- list("unburned" = "unburned", "low severity" = "<50", "high severity" = ">50")
levels(iagsum$YEAR) <- list("2016" = "16", "2018" = "18", "2019" = "19", "2020" = "20")

rawiagprop <- ggplot(iagsum, aes(x = as.integer(as.character(YEAR)), y = prop.IAG, group = burn2, color = burn2) ) + 
		geom_line(position = position_dodge(0.2))+
		geom_point(size = 3.2, position = position_dodge(0.2)) + # Add points
  geom_errorbar(aes(ymin = prop.IAG - se, ymax = prop.IAG + se), width = .5, alpha = 0.5, size = 1, position = position_dodge(0.2))+
		labs(x = NULL, y = "Proportion IAG") + # Label axes
		theme_bw(base_size = 14) + # Make black and white,
								# increase base text size
			 theme(panel.grid.major.x = element_blank())+ # Remove y gridlines
			         scale_color_manual(values = c("#52796f", "#dda15e", "#bc6c25"))+
  theme(legend.position = "bottom")+ 
  theme(legend.title = element_blank()) +
  ggtitle("IAG-to-total cover") +
  geom_vline(xintercept = 2017, colour = "dark grey", lty = 2)+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())


#ggsave("propIAGraw.svg")
```


```{r fig.height=8, fig.width=12}
legend <- get_legend(rawvd)

rawvd <- rawvd + theme(legend.position="none") 

rawplots <- ggarrange(rawvd, rawbt, rawbj, rawiagprop, ncol = 4, nrow = 1, common.legend = TRUE, legend="bottom")

contrastplots <- ggarrange(vcontrast, btcontrast, bjcontrast, iagcontrast, ncol = 4, nrow = 1, common.legend = TRUE, legend="bottom")

allplots <- ggarrange(rawplots, contrastplots, nrow = 2)
  
allplots


# ggsave("all_IAG_models.svg")
# ggsave("annual_Grass_cover_through_time_raw.png")
```

