---
title: "upper beaver"
author: "Claire Tortorelli"
date: "January 2, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyr)
library(ordinal)
library(ggplot2)
library(MASS)
library(brant) 

```

organize data
```{r read in data}

ub_veg <- read.csv(here("raw-data","ubdata_16-20_species_average.csv"))

ub_veg$CANOPY <- ub_veg$JUOC + ub_veg$PIPO + ub_veg$PSME

#extract columns for analysis
sub = ub_veg[c("PLOT","YEAR","BURN17", "VEDU", "BRJA", "BRTE", "CANOPY")]
#change year to factor
sub$YEAR <- as.factor(sub$YEAR)

#add IAG col
sub$IAGperc <- sub$BRJA + sub$BRTE  + sub$VEDU


```

```{r exploring vedu effect on burn}
#https://marissabarlaz.github.io/portfolio/ols/
#helpful webpage! ^^


#remove control plots
orddf <- subset(sub, BURN17 != "c_unburned")
#remove all data collected after the burn
orddf <- subset(orddf, YEAR == 16)

#change burn level to a factor
orddf$BURN17 <- factor(as.character(orddf$BURN17), levels = c("unburned","<50",">50"))
levels(orddf$BURN17) <- list(unburned="unburned", low_sev="<50", high_sev=">50")

qplot(x = IAGperc, y = BURN17, data = orddf)

```
### Model response
Ordinal regression with cumulative link model - because our response is an ordered variable (unburned < low sev < high sev)
```{r}
#helpful documentation for clm & ordinal regressions- http://people.vcu.edu/~dbandyop/BIOS625/CLM_R.pdf

ord1 <-clm(BURN17 ~ IAGperc + CANOPY, data = orddf) 

summary(ord1) #Wald tests
#Hessian is a measure of how identifiable the model is; large values, say larger than 1e4 indicate that the model may be ill defined. 
# - this models Hessian is e^3

#liklihood ratio test
drop1(ord1, test = "Chi") 

#test the effect of the variables while ignoring the effect of the other variable
ord2 <- clm(BURN17 ~ 1, data = orddf)
add1(ord2, scope = ~ IAGperc + CANOPY, test = "Chi")

#liklihood ratio test
library(lmtest)
lrtest(ord1, ord2) 
#CANOPY improves to model fit


#plot profiles
plot(profile(ord1))

```
Check assumptions
```{r}
#checking clm assumptions. IAG is sig. with the nominal-test - that the slope between each pair of outcomes across 2 response levels is the same..
nominal_test(ord1)
scale_test(ord1)

#check correlation of canopy and IAGall
cor(orddf$IAGperc, orddf$CANOPY) #pretty low, -.33
```

Testing the proportional odds/parallel regression assumption
```{r}
#remodling with polr to test parrall regression assumption
ord3 <- polr(BURN17 ~ IAGperc + CANOPY, data = orddf, Hess = TRUE)
summary(ord3)

#test parallel regression assumption with brant function
brant(ord3) 
# parallel regression assumption has been violated!
```


Data do not appear to meet the assumptions of ordinal regression or partial proportional odds. Using a generalized ordinal model as it relaxes the proportional odds assumption

```{r}
library(oglmx)
#estimate ordered logit with heteroskedasticity
gord1 <- oglmx(BURN17 ~ IAGperc + CANOPY, ~ IAGperc, data= orddf, link = "logit",  delta = 1, threshparam = NULL, constantMEAN = FALSE, constantSD = FALSE)
summary(gord1)
margins.oglmx(gord1, ascontinuous = TRUE)


#compare models with liklihood ratio test

lrtest(ord1, gord1)

#the simpler model (standard ordered logit) is a better fit than the heteroskadastic model
```

Using multinomial model as it does not assume normality, linearity, or homoscedasticity

Multinomial model
```{r}
library(foreign)
library(nnet)


m1 <- multinom(BURN17 ~ IAGperc + CANOPY, data = orddf) # does not assume linearity, normality, or homoskedasticity, just independence
summary(m1)
confint(m1)

z <- summary(m1)$coefficients/summary(m1)$standard.errors
z

# 2-tailed z test
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p



#multinomial model is a better fit than ordinal regression and assumptions are met
```

exploring results
```{r}
## extract the coefficients from the model and exponentiate
exp(coef(m1))

#The relative risk ratio for a one-unit increase in the variable IAGperc is .961 for being <50% burned vs. unburned

head(pp <- fitted(m1))

```


```{r}
#compare multinomial model to ordinal model
lrtest(m1,ord1)
summary(m1) #AIC 59
summary(ord1) #AIC 62

```
Decided to use a multinomial model because it fits the data best and the data meet all assumptions. 

Instead of calculating probabilities of transitions, we'll calculate the probability of burning in each category relative to the unburned category


#examine the changes in predicted probability associated with canopy held constant
```{r}
dfplot <- data.frame(IAGperc = c(10,25,30,40,50,60), CANOPY = mean(orddf$CANOPY))
preds <- predict(m1, newdata = dfplot, "probs")

```

### plot model
Extract predictions for plotting
```{r}
library(effects)

#extract fits and 95% cis
fit.eff <- Effect("IAGperc", m1, xlevels = 200)
fit.df <- data.frame(fit.eff$model.matrix, fit.eff$prob, fit.eff$lower.prob, 
           fit.eff$upper.prob)

#create predictor df for plotting
fit.df2 <- gather(fit.df, "burncat", "prob", c(prob.unburned, prob.low_sev, prob.high_sev))

#unburned
fit.df.ub <- subset(fit.df2[,c(1:3,6,9:10)], burncat == "prob.unburned")
#low severity
fit.df.ls <- subset(fit.df2[,c(1:2,4,7,9:10)], burncat == "prob.low_sev")
#high severity
fit.df.hs <- subset(fit.df2[,c(1:2,5,8:10)], burncat == "prob.high_sev")
colnames(fit.df.ub)[3:4] <- c("l.ci", "u.ci")
colnames(fit.df.ls)[3:4] <- c("l.ci", "u.ci")
colnames(fit.df.hs)[3:4] <- c("l.ci", "u.ci")

multdf <- rbind(fit.df.ub, fit.df.ls, fit.df.hs)
multdf$burncat <- factor(multdf$burncat, levels = c("prob.unburned", "prob.low_sev", "prob.high_sev"), labels = c("unburned", "low severity", "high severity"))

newdf <- data.frame(
  IAG_cover = c(fit.df$IAGperc,fit.df$IAGperc,fit.df$IAGperc),
  prob = c(fit.df$prob.unburned, fit.df$prob.low_sev, fit.df$prob.high_sev),
  l.ci = c(fit.df$L.prob.unburned, fit.df$L.prob.low_sev, fit.df$L.prob.high_sev),
  u.ci = c(fit.df$U.prob.unburned, fit.df$U.prob.low_sev, fit.df$U.prob.high_sev),
  burncat = factor(rep(c("unburned", "low severity", "high severity"), each=200), levels = c("unburned", "low severity", "high severity")))
```

Plot standard ordinal regression
```{r fig.height=4, fig.width=4.25}
#probability that a site burned in one of these three classes is related to pre-fire IAG cover, after accounting for canopy cover

library(devtools)
library(ggeffects)

ggplot(newdf, aes(x = IAG_cover, y = prob)) + 
    geom_ribbon(aes(ymin=l.ci, ymax=u.ci, fill = burncat), alpha = 0.3) +
  geom_line(aes(color = burncat, group = burncat), se = FALSE, size = 1.2) +
  scale_color_manual(values = c("#52796f", "#dda15e", "#bc6c25"))+
  scale_fill_manual(values = c("grey", "grey", "grey"))+
  theme_bw(base_size = 13)+
  theme(legend.title = element_blank(), legend.position = "bottom")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  xlab("IAG cover (%)")+
  ylab("Probability")

ggsave("multinomial_by_IAG.png")
```




